Hydra battles in Coq

### 2020-10-01 doc/hydras.tex
minor modifications

Instance Omega2 : ON  lt compare.
Proof.
  split.
  - apply lt_strorder.
  - apply lt_wf.
  - apply compare_correct.
Qed.
\end{Coqsrc}

@@

\begin{Coqsrc}
(** multiplication of an ordinal and a natural number *)

@@

(** multiplication of  a natural number and an ordinal *)

@@

\subsection{Cantor normal form}

The ordinal \(\epsilon_0\) is the least ordinal number that satisfies 
the equation \(\alpha = \omega^\alpha\), where \(\omega\) is 
the least infinite ordinal. Thus, we can consider \(\epsilon_0\) as an
\emph{infinite} \(\omega\)-tower.

\index{Maths!Cantor normal form}

Any ordinal strictly less that \(\epsilon_0\) 
can be represented by a unique  \emph{Cantor normal form}, 
that is, an expression  which is either  the ordinal \(0\) or 
a sum  \(\omega^{\alpha_1} \times n_1 + \omega^{\alpha_2} \times n_2 + 
  \dots + \omega^{\alpha_p} \times n_p\) where all the \(\alpha_i\) 
are ordinals in Cantor  normal form, \(\alpha_1 > \alpha_2 > \alpha_p\), 
and all the \(n_i\) are positive integers.

@@

\paragraph{Remark}
For simplicity's sake, we chosed to forbid  expressions of the form $\omega^\alpha\times 0 + \beta$. Thus, the contruction (\texttt{ocons $\alpha$ $n$ $\beta$} is intented to represent the
ordinal $\omega^\alpha\times(n+1)+\beta$ and not $\omega^\alpha\times n+\beta$.
In a future version, we should replace  the type \texttt{nat} with \texttt{positive} in \texttt{T1}'s 
definition. But this replacement would take a lot of time \dots{}

@@

Some abbreviations may help to write more consisely complex ordinal terms.

@@

Notation "'F' n" := (fin n) (at level 10) : t1_scope.

@@

The ordinal $\epsilon_0$, although not represented by a finite term in Cantor normal form, is approximed by the sequence of $\omega$-towers.

@@

\begin{Coqsrc}
Theorem lt_irrefl  alpha :  ~ lt alpha  alpha.

@@

\subsubsection{A predicate for characterizing normal forms}
\label{sec:orgheadline74}
\label{sec:orgheadline75}
We note that our data-type \texttt{T1} allows us to write expressions that
are not properly in Cantor normal form as specified in Section \ref{sec:epsilon0-intro}.
For instance, consider the following term of type  \texttt{T1}.

@@

But this style leads to clumsy statements, and many sub-goals in interactive proofs (although often solved with \texttt{auto} or \texttt{eauto}).

@@

\begin{Coqsrc}
Definition LT := restrict nf lt.
Infix "o<" := LT : t1_scope.

Definition LE := restrict nf le.
Notation "alpha o<= beta" := (LE alpha beta) : t1_scope.
\end{Coqsrc}

\begin{todo}
Infix notation for inequalities on ordinals have changed to \texttt{o<} and \texttt{o<=}.  This document is not updated yet, but the \coq{} scripts are correct.
\end{todo}

@@

\begin{Coqsrc}
Lemma LT_one : forall alpha, alpha o< one -> alpha = zero.
\end{Coqsrc}

@@

\begin{todo}
Infix notation for inequalities on type \texttt{T1} have changed to \texttt{t1<} and \texttt{t1<=}.  This document is not updated yet, but the \coq{} scripts are correct.
\end{todo}

@@

Then, the ordinal  $\epsilon_0$ is defined as the limit of the sequence of all finite towers (a kind of infinite tower).

>>>>

Instance Omega2: ON lt compare.
Proof.
  split.
  - apply lt_strorder.
  - apply lt_wf.
  - apply compare_correct.
Qed.
\end{Coqsrc}

@@

\begin{Coqsrc}
(** multiplication of an ordinal by a natural number *)

@@

(** multiplication of  a natural number by an ordinal *)

@@

\subsection{Cantor normal form}
\index{Maths!Cantor normal form}

The ordinal \(\epsilon_0\) is the least ordinal number that satisfies 
the equation \(\alpha = \omega^\alpha\), where \(\omega\) is 
the least infinite ordinal. Thus, we can consider \(\epsilon_0\) as an
\emph{infinite} \(\omega\)-tower.
Nevertheless, 
any ordinal strictly less that \(\epsilon_0\) 
can be finitely represented by a unique  \emph{Cantor normal form}, 
that is, an expression  which is either  the ordinal \(0\) or 
a sum  \(\omega^{\alpha_1} \times n_1 + \omega^{\alpha_2} \times n_2 + 
  \dots + \omega^{\alpha_p} \times n_p\) where all the \(\alpha_i\) 
are ordinals in Cantor  normal form, \(\alpha_1 > \alpha_2 > \alpha_p\), 
and all the \(n_i\) are positive integers.

@@

\paragraph{Remark}
For simplicity's sake, we chosed to forbid  expressions of the form $\omega^\alpha\times 0 + \beta$. Thus, the contruction (\texttt{ocons $\alpha$ $n$ $\beta$}) is intented to represent the
ordinal $\omega^\alpha\times(n+1)+\beta$ and not $\omega^\alpha\times n+\beta$.
In a future version, we should replace  the type \texttt{nat} with \texttt{positive} in \texttt{T1}'s 
definition. But this replacement would take a lot of time \dots{}

@@

Some abbreviations may help to write more concisely complex ordinal terms.

@@

@@

The ordinal $\epsilon_0$, although not represented by a finite term in Cantor normal form, is approximed by the sequence of $\omega$-towers (see also Sect~\vref{sect:epsilon0-as-limit} ).

@@

\begin{Coqsrc}
Theorem lt_irrefl (alpha: T1):  ~ lt alpha alpha.

@@

\subsubsection{A predicate for characterizing normal forms}
\label{sec:orgheadline74}
\label{sec:orgheadline75}
Our data-type \texttt{T1} allows us to write expressions that
are not properly in Cantor normal form as specified in Section \ref{sec:epsilon0-intro}.
For instance, consider the following term of type  \texttt{T1}.

@@

But this style leads to clumsy statements, and generates too many sub-goals in interactive proofs (although often solved with \texttt{auto} or \texttt{eauto}).

@@

\begin{Coqsrc}
Definition LT := restrict nf lt.
Infix "t1<" := LT : t1_scope.

Definition LE := restrict nf le.
Infix "t1<=" := LE : t1_scope.
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma LT_one : forall alpha, alpha t1< one -> alpha = zero.
\end{Coqsrc}

@@

@@

\label{sect:epsilon0-as-limit}
Then, the ordinal  $\epsilon_0$ is defined as the limit of the sequence of all finite towers (a kind of infinite tower).
